---
title: "Module 2 Final Project — Milestone 1: EDA"
author: "Abby Bridgers and Justin Ehringhaus"
date: "July 24, 2022"
output:
  html_document: default
bibliography: "references.bib"
nocite: '@*'
---
## AB test xxx
---

Test comment.

```{r setup, include=FALSE}
# repo: https://github.com/ab6723/aly6040_V02
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
p_load(tidyverse)  # usual suite of packages
p_load(FSA)        # view first and last rows of a data frame: headtail()
```

---

## Data Preparation

```{r}
wvs <- read_csv("WVS_Cross-National_Wave_7_csv_v4_0.csv")
headtail(wvs, n = 3, which = c(1:3)) # first/last 3 rows, first/last 3 columns
```

- Explore the variables, data types, values, etc... 

## Analysis & Interpretation

- Calculate appropriate summary statistics
- Create appropriate graphs that will give you insights into the data.

Include the answers to following questions:

- What did you do with the data in the context of exploration?
- How many entries are in the dataset?
- Was there missing data? Duplications? How clean was the data?
- Were their outliers or suspicious data?
- What did you do to clean the data?
- What did you find? What intrigued you about the data? Why does that matter?
- What would your proposed next steps be?
- What business questions do you plan to answer with your data mining?
