---
title: "Module 2 Final Project — Milestone 1: EDA"
author: "Abby Bridgers and Justin Ehringhaus"
date: "July 24, 2022"
output:
  html_document: default
  # github_document: default
bibliography: "references.bib"
nocite: '@*'
---

---

```{r setup, include=FALSE}
# repo: https://github.com/ab6723/aly6040_V02
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
p_load(tidyverse)  # usual suite of packages
p_load(ggthemes)   # extra themes
p_load(hrbrthemes) # more extra themes
p_load(skimr)      # alternative to summary(), skims dataset: skim()
p_load(VIM)        # visualization of missing values: aggr()
p_load(corrplot)   # visualization of a correlation matrix: corrplot()
p_load(maps)       # world map
```

---

## Introduction

The World Values Survey (WVS) aims to understand the values, beliefs, and norms of people all over the world, comparatively and longitudinally. Operating in more than 120 countries and conducted every five years, the WVS uses a common questionnaire and household interview format to investigate human beliefs. The data is made available in waves, where the most recent Wave 7 represents interviews conducted between 2017-2022. Time series data including all waves pooled between 1981-2022 is also available.

While examining the pooled dataset would certainly be interesting to answer questions about the changes worldwide in people’s values, beliefs, and norms, we are opting to limit the scope of our project to recent years (Wave 7: 2017-2022). Already, the size and scope of this dataset represents a big step for our group’s budding data analysts, as we have never yet worked with anything so faceted. The Wave 7 dataset is wide, containing 552 variables and 87,822 rows, and thus it will be necessary to subset those variables that interest us the most.

Our team (Abby Bridgers and Justin Ehringhaus) are interested in how universal certain values are as well as how values differentiate across the globe. Abby is hoping to learn specifically about values on immigration, and Justin is hoping to learn specifically about social views and ethical values. 

##### What did you do with the data in the context of exploration?

We created subsets with our data to explore and analyze from two perspectives. This allowed us to form separate viewpoints and then pool our knowledge to better understand the steps necessary to prepare the data for further analysis.

##### How many entries are in the dataset?

The WVS Wave 7 dataset contains 552 variables and 87,822 rows. We reduced the number of attributes by about 90%. More details follow.

##### Missing data? Duplicates? Cleanliness? Outliers?

While there was missing data found in our dataset, we elected to refrain from cleaning the data because doing so may unintentionally falsify the results of the WVS survey. In a typical data cleaning process, outliers are removed to concentrate on the central tendency of the data. Missing data is removed and/or standardized for numerical analysis. Since our data is ordinal, and we are interested in including and analyzing the data holistically, we did not feel that it would be appropriate "clean" the data beyond changing column names, selecting attributes of interest, and ensuring that there was not a prohibitory amount of NA values found in our subsets of interest. 

---

## Data Preparation

```{r message=FALSE, warning=FALSE}
# importing the entire dataset
wvs <- read_csv("../WVS_Cross-National_Wave_7_csv_v4_0.csv")

# checking first few rows and columns for import success
head(wvs)[,1:5]
```

```{r}
wvs_subset <- 
  wvs %>% 
  select(
    # --------------------------- DEMOGRAPHICS
    Country = B_COUNTRY_ALPHA,
    Longitude = O1_LONGITUDE,
    Latitude = O2_LATITUDE,
    Settlement.type = H_SETTLEMENT,
    Country.and.year = S025,
    Town.size = G_TOWNSIZE2,
    Age = Q262,
    Income.Group = Q288,
    Ethnic.Group = Q290, # see WVS_codebook.pdf for Q290 coding info
    Immigrant = Q263,
    Religion = Q289,
    Marital.Status = Q273,
    Education = Q275,
    Number.Children = Q274,
    Happiness = Q46,
    Health = Q47,
    # --------------------------- POLITICAL PARTICIPATION / CONFIDENCE IN GOVERNMENT
    votes.locally = Q221,
    votes.nationally = Q222,
    confidence.elections = Q76,
    confidence.courts = Q70,
    confidence.UN = Q83,
    environment.vs.econgrow = Q111,
    # --------------------------- RELATIONSHIP BETWEEN GOVERNMENT AND CITIZENS
    cheating.taxes = Q180,
    gov.video.surveillance = Q196,
    gov.email.monitoring = Q197,
    gov.collecting.info = Q198,
    # --------------------------- ETHICAL VALUES ---------------------------
    terrorism = Q192,
    death.penalty = Q195,
    suicide = Q187,
    beating.wife = Q189,
    beating.children = Q190,
    # --------------------------- SOCIAL VIEWS ---------------------------
    homosexuality = Q182,
    prostitution = Q183,
    abortion = Q184,
    divorce = Q185,
    casual.sex = Q193,
    sex.before.marriage = Q186,
    # --------------------------- CAREER VALUES ---------------------------
    importance.leisure.time = Q3,
    importance.work = Q5,
    # --------------------------- IMMIGRATION ---------------------------
    job.scarc.prioritizes.nonimm = Q34,
    imm.fills.useful.jobs = Q122,
    imm.strengthens.cultural.div = Q123,
    imm.increases.crime.rate = Q124,
    imm.gives.political.asylum = Q125,
    imm.increases.terrorism.risk = Q126,
    imm.helps.poor = Q127,
    imm.increases.unemployment = Q128,
    imm.brings.social.conflict = Q129,
    imm.policy.preference = Q130
    )
```

---

## Analysis & Interpretation

```{r}
myskim <- skim(wvs_subset)
myskim
```

The subset of the dataset contains `r nrow(wvs_subset)` rows and `r ncol(wvs_subset)` columns.

Most variables are numeric; only `Country` is classified as a character vector. `Longitude` and `Latitude` are missing the most data with a completion rate of `r myskim$complete_rate[2]`, but overall the mean completion rate is `r mean(myskim$complete_rate)`, which signifies that missing data is not too much of an issue for the particular variables under consideration.

As many of the variables are Likert scales, the values are ordinal ranging from 1 to 5, for example, and thus the descriptive statistics generated such as mean, sd, quantiles, and histograms give good indications of the distribution and the characteristics of the data. However, certain variables such as `Ethnic.Group`, `Country.and.year`, `Longitude`, and `Latitude` contains nominal, numeric data, and thus the descriptive statistics generated should be ignored.

```{r}
# unique values and associated counts for categorical data only
wvs_subset_categorical <- 
  wvs_subset %>% 
  select(-Longitude,
         -Latitude)

lapply(wvs_subset_categorical, table)
```

The unique values and associated counts of the categorical data above reveals the dataset is clean. Thankfully, data collection and data entry efforts conducted by the World Values Survey appear to be rigorous, and thus there do not appear to be instances of inappropriately assigned values.

That said, one question for consideration is whether to consider abnormal or atypical experiences as outlier information. For example, if we examine the `Number.Children` variable, we can see the minimum value is `r min(wvs_subset$Number.Children, na.rm = TRUE)` and the maximum value is `r max(wvs_subset$Number.Children, na.rm = TRUE)`. Clearly, that is a high maximum, but if we assume the value was not entered incorrectly, then this is an actual facet of the human experience and should not be treated as an outlier. Removing such data points from our analysis would, in fact, skew our perception of the human experience, and thus for this particular dataset we will opt not to remove any pieces of information with the assumption that data has rigorously been collected and verified by research professionals.

```{r}
# ordinal, numeric data only 
wvs_subset_ordinal <- 
  wvs_subset %>% 
  select(-Longitude,
         -Latitude,
         -Country, 
         -Ethnic.Group, 
         -Country.and.year,
         -Religion)

# assessing correlations between ordinal, numeric data
# only including data where pairwise observations are complete (i.e., not including missing data)
cor <- cor(wvs_subset_ordinal, use = "pairwise.complete.obs")
corrplot(cor, 
         method = "circle", 
         insig = "blank", 
         diag = FALSE,
         tl.cex = 0.5) %>% 
    corrRect(name = c('Settlement.type', 
                      'votes.locally', 
                      'cheating.taxes', 
                      'terrorism', 
                      'homosexuality',
                      'importance.leisure.time',
                      'job.scarc.prioritizes.nonimm',
                      'imm.policy.preference'))
```

The visualization of a correlation matrix above indicates the extent to which each variable, compared to one another, correlates or relates. Darker blue colors represent stronger positive correlations, and darker orange colors represent stronger negative correlations. For visualization purposes, we have bucketed the variables into four groups, albeit it must be noted that by doing so we risk imposing structure where it does not, or should not, exist: 

- The first bucket contains variables relating to demographics
- The second relates to political participation / confidence in government
- The third relates to the relationship between government and citizens
- The fourth relates to ethical values
- The fifth relates to social views
- The sixth relates to career values
- The seventh relates to immigration

There are endless conversations that could be held based off this visualization of a correlation matrix. Some key insights are as follows:

- All variables in the fifth bucket (social views) are strongly correlated. This makes sense, as those with similar social views are generally split along the lines of liberal versus conservative thought. For example, it makes sense that those who hold conservative views on sexuality also hold conservative views on marriage, career, and government.
- Education is positively correlated with liberal social views (i.e., higher education = more liberal social views).
- Number of children is negatively correlated with liberal social views (i.e., more children = more conservative social views).
- Education is negatively correlated with number of children (i.e., higher education = less children).
- Voting locally or nationally is negatively correlated with age (i.e., older people participate more in politics).
- Views on terrorism positively correlate with views on domestic abuse (i.e., those who view terrorism favorably also view domestic abuse favorably).
- There is a moderate correlation between two questions on immigration. In the immigration questions, respondents were asked to answered their degree of agreement with statements. The moderately positively correlated statements are "Immigration in my country increases crime rate" and "immigration in my country brings social conflict." These questions are answered on the same ordinal scale of 3 options: "disagree", "hard to say", and "agree". 
- There is a strong negative correlation between settlement type and town size which are both on a 5-point scale. A response of "1" to the town size question means the respondent lives in a town of under 5,000 people. A response of "1" to the settlement type question means the respondent lives in the capital city of their country. Therefore, there is a strong inverse relationship between these two attributes.

The above observations from the correlation matrix are just several of many possible avenues for exploration. By understanding how different variables correlate, we can begin to explore specific facets of the dataset in more detail.

```{r}
# Creates colored points (gradient) based on particular categorical numeric values in dataframe.
# For display in a map (see examples below)
map_color_range <- function(tibl, col, 
                            min_color = "green", max_color = "red",
                            shape = 1, size = 0.4) {
  min_value <- min(col, na.rm = TRUE)
  max_value <- max(col, na.rm = TRUE)
  # vector of color gradients with a length of `max_value`
  colors <- colorRampPalette(c(min_color, max_color))(max_value)
  # executes points() for each value between min and max
  for (i in min_value:max_value) {
    tibl_subset_col <- 
      tibl %>% 
      filter(col == i) 
    points(tibl_subset_col$Longitude, 
           tibl_subset_col$Latitude, 
           col = alpha(colors[i], 0.1), # colors from color ramp above w/ transparency
           pch = shape, 
           cex = size)
  }
}

map("world")
title(main = "Happiest (Green) vs. Unhappiest (Red)")
map_color_range(wvs_subset, wvs_subset$Happiness)
```

Another way of exploring the data is to investigate geospatial relationships using longitude and latitude data paired with variables of interest. For example, the visualization above reveals respondents happiness levels from "very happy" (greener points) to "not at all happy" (redder points). Pockets of unhappiness appear clustered in South America, parts of Africa, and parts of Europe and Asia. Generally, the United States and Mexico appear happier.

Geospatial data can reveal many interesting trends. As another example, we can visualize beliefs on homosexuality, where greener colors represent those who believe homosexuality is "always justifiable" and redder colors represent those who believe homosexuality is "never justifiable." Eastern Europe and parts of Africa appear most opposed to homosexuality.

```{r}
map("world")
title(main = "Believes Homosexuality is Justifiable (Green) vs. Never Justifiable (Red)")
map_color_range(wvs_subset, wvs_subset$homosexuality, 
                min_color = "red", max_color = "green")
```

The relationship between geography and health looks remarkably similar to that of geography and happiness. As is the case with happiness, the US and Mexico appear most healthy, with strong clusters of poor health in Africa, South America, and Europe, to name a few. Southeast Asia appears to be one of the most diverse areas in terms of health, with a great deal of both blue and yellow shown on the map. 

```{r}
map("world")
title(main = "Good Health (Yellow) vs. Poor Health (Blue)")
map_color_range(wvs_subset, wvs_subset$Health, 
                min_color = "yellow", max_color = "blue")
```

As we could not include the `Religion` variable in our correlation matrix above as it is a nominal rather than ordinal variable, it will be interesting to explore religion in relation to `Number.Children`, for example. The visualization below reveals some interesting trends in religious affiliation and family size:

- Muslims tends to have the most children proportionally and in the highest numbers.
- Those without a religion generally have fewer children.
- Only Roman Catholics and Protestants have 23+ children, which signifies there may be only one instance (one individual) accounting for this particular data. As discussed previously, we will not consider this an outlier as this is a fact of the human experience. Removing such data would skew our perception of what occurs in the world.

```{r}
Religion.names <- c('No Religion', 'Roman Catholic', 'Protestant', 'Orthodox',
                    'Jew', 'Muslim', 'Hindu', 'Buddhist', 'Other Christian', 'Other', 'NA')

wvs_subset %>% 
  ggplot +
  aes(x = factor(Number.Children), fill = factor(Religion)) +
  geom_bar(position = "fill") +
  scale_fill_discrete(name = "Religious Denomination",
                      labels = Religion.names) +
  xlab("Number of Children") +
  ylab("Proportion") +
  ggtitle("Religious Affiliation and Number of Children:\nWhich Religions Proportionally Have the Most Children?") +
  theme_tufte()
```

```{r}
# basic bar chart
ggplot(wvs_subset, aes(x=confidence.courts, y=Age)) + 
  geom_bar(stat = "identity")
```

In the above barchart, confidence in the courts is measured. Confidence ranges from "A great deal" (1) to "Not at all" (4). Most respondents chose 2, meaning they have "Quite a lot" of confidence in the courts. 

```{r}
# histogram of age of Indonesian respondents 
p <- wvs_subset %>%
  filter( Country=="IDN" ) %>%
  ggplot( aes(x=Age)) +
    geom_histogram( binwidth=3, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
    ggtitle("Age of Indonesian Respondents") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=20)
    )
p 
```

In the above histogram, the age of Indonesian respondents is shown. Very few respondents are over the age of 75. The data is clustered around age 30, with the majority of respondents being between ages 25 and 50. The average age in Indonesia is around 30, which aligns with the average age worldwide. Conversely, Germany has a high average age of 48. In the following histogram, I find the age of German respondents to see if the data is clustered around a higher average age than 30 given the much higher country average.

```{r}
# histogram of age of German respondents 
p <- wvs_subset %>%
  filter( Country=="DEU" ) %>%
  ggplot( aes(x=Age)) +
    geom_histogram( binwidth=3, fill="blue4", color="#e9ecef", alpha=0.9) +
    ggtitle("Age of German Respondents") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=20)
    )
p 
```

As expected, the above histogram shows that the average age of German respondents is closer to 50. There may be a difference in viewpoints among those age 30 and 50, making age an important factor for continued consideration in our project.

```{r}
# age overlaid with income in prop bar chart
Income.details <- c('Lower Step', 'Second Step', 'Third Step', 'Fourth Step', 'Fifth Step', 'Sixth Step', 'Seventh Step', 'Eigth Step', 'Nineth Step', 'Tenth Step', 'Unknown', 'No Answer', 'Not Asked', 'NA')

wvs_subset %>% 
  ggplot +
  aes(x = Age, fill = factor(Income.Group)) +
  geom_bar(position = "fill") +
  scale_fill_discrete(name = "Income Group",
                      labels = Income.details) +
  xlab("Age") +
  ylab("Proportion") +
  ggtitle("Income Group Proportion By Age") +
  theme_hc()
```

The above proportional stacked bar graph allows for visualizing differences in income by age. Respondents were shown cards with their country's income broken down into ten buckets. Since this varies heavily by country, the Wave 7 results do not include any specific numbers. Instead, income is purely comparative. In this chart, several middle income groups appear to fall with increasing age; third through sixth steps noticably decrease from young respondents through middle age respondents to elderly respondents. The highest income, tenth step, appears relatively constant across ages, with a noticable increase in variance from age to age among those in and beyond their 80s. The most common income group appears to be the fifth step, which may mean that most people consider themselves middle-class in the context of their home country. 

## Conclusions

##### What did you find? What intrigued you about the data? Why does that matter?

- We were impressed by the cleanliness of the dataset after completing initial exploratory analysis. 
- We found several correlations, including social, family, immigration, and demographic information. 
- For instance, we found some interesting insights around education. Education is positively correlated with liberal social views, meaning that respondents with post-secondary degrees have more liberal social views. Education is negatively correlated with quantity of children, meaning that respondents with post-secondary degrees have fewer children. 
- To offer a second example, we found moderate correlations with several immigration variables. Respondents with negative views towards immigration tend to feel negatively across variables. 
- We found interesting patterns by looking at the scale of a variable on a map of the world. For instance, happiness appeared highly variable by geography. for instance, Russia, the US, Mexico, Argentina, and the Philippines all appear very green, indicating high rates of happiness in those countries. In contrast, Zimbabwe, Egypt, Iran, Bolivia, and certain parts of China are very red, indicating high rates of unhappiness in those countries. 
- In addition, we explored religious affiliation and number of children using a proportional bar chart. We found that the respondent with 24 children is Protestant. In addition, from this chart, we were able to see the religious makeup of respondents; the plurality of respondents are Hindus.
- We also explored income group proportion by age, finding that the highest income group was fairly equally represented across ages, while other income groups, such as the lowest step and the third-lowest step, increased in proportion with age. 

These preliminary findings matter because they shed light on differences in identity, which can be contrasted to social, political, and ethical viewpoints to deduce relationships between identity and belief.

##### What would your proposed next steps be?

- Share learning and techniques with classmates.
- Integrate feedback from Professor Hodeghatta to improve our analysis and streamline our path forward.
- Uncover answers to our key questions by applying data mining techniques including decision trees, clusters, association mining, and linear discriminant analysis.
- Continue to write about our findings and comment throughout the code.

##### What business questions do you plan to answer with your data mining?

- Predicting the happiness and health of an individual.
- Predicting the religious affiliation of an individual.
- Predicting the immigration views, social views, and ethical values of an individual.
- The relationship between income (Q288), happiness (Q46), and health (Q47).
- The relationship between marital status (Q273) and political engagement (Q209 - Q222).
- The relationship between religious denomination (Q289) and number of children (Q274).
- The differences in ethical values (Q177-Q195) by ethnic group (Q290).
- The relationship between immigration stance (122:130) and age (262).
- The relationship between education completed (Q275) and the importance of leisure time (Q3).

## Works Cited