---
title: "Module 6 Final Project — Final Report"
author: Abby Bridgers and Justin Ehringhaus
date: Last edited `r format(Sys.time(), '%B %d, %Y')` at `r format(Sys.time(), '%H:%M')`
output:
  github_document: 
    toc: true
    toc_depth: 5
bibliography: "references.bib"
nocite: '@*'
---

---

```{r setup, include=FALSE}
# repo: https://github.com/ab6723/aly6040_V02
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
p_load(tidyverse)  # usual suite of packages
p_load(ggthemes)   # extra themes
p_load(hrbrthemes) # more extra themes
p_load(skimr)      # alternative to summary(), skims dataset: skim()
p_load(VIM)        # visualization of missing values: aggr()
p_load(corrplot)   # visualization of a correlation matrix: corrplot()
p_load(maps)       # world map
p_load(rattle)
p_load(rpart, rpart.plot)
p_load(caret)
```

---

## Introduction

##### Overview

The World Values Survey (WVS) aims to understand the values, beliefs, and norms of people all over the world, comparatively and longitudinally. Operating in more than 120 countries and conducted every five years, the WVS uses a common questionnaire and a household interview format to investigate human beliefs. The data is made available in waves. Time series data including all waves pooled between 1981-2022 is also available.

While examining the pooled dataset would certainly be interesting to answer questions about changes in worldwide values, beliefs, and norms, we are opting to limit the scope of our project to recent years (Wave 7: 2017-2022). Already, the size and scope of this dataset represents a big step for our group's budding data analysts, as we have never yet worked with anything so large and so faceted. The Wave 7 dataset contains 552 variables and 87,822 rows. We have reduced the number of attributes by about 90%. More details to follow.

##### Methods and Business Questions

Our team is interested in exploring values on religion, immigration, ethics, and governmental-related topics. We analyze particular subsets of the dataset using exploratory and predictive techniques suited to the size, scope, and types of data under consideration. For example, we chose to incorporate geospatial visualization techniques as our dataset relates to the world, and we chose to orient our predictive modeling around decision trees because of our restriction to ordinal (Likert) data. Our analyses are inspired by the four business questions listed below:

- #1: Predicting whether or not an individual is an immigrant
- #2: Predicting whether an individual views teaching tolerance as important based on their beliefs regarding immigration 
- #3: Predicting the religious denomination of an individual
- #4: Predicting an individual's confidence in elections

---

## Data Preparation

```{r message=FALSE, warning=FALSE}
# importing the entire dataset
wvs <- read_csv("../WVS_Cross-National_Wave_7_csv_v4_0.csv")

# checking first few rows and columns for import success
head(wvs)[,1:5]
```

```{r}
wvs_subset <- 
  wvs %>% 
  select(
    # --------------------------- DEMOGRAPHICS
    Country = B_COUNTRY_ALPHA,
    Longitude = O1_LONGITUDE,
    Latitude = O2_LATITUDE,
    Settlement.type = H_SETTLEMENT,
    Country.and.year = S025,
    Town.size = G_TOWNSIZE2,
    Age = Q262,
    Income.Group = Q288,
    Ethnic.Group = Q290, # see WVS_codebook.pdf for Q290 coding info
    Immigrant = Q263,
    Religion = Q289,
    Marital.Status = Q273,
    Education = Q275,
    Number.Children = Q274,
    Happiness = Q46,
    Health = Q47,
    # --------------------------- POLITICAL PARTICIPATION / CONFIDENCE IN GOVERNMENT
    votes.locally = Q221,
    votes.nationally = Q222,
    confidence.elections = Q76,
    confidence.courts = Q70,
    confidence.UN = Q83,
    environment.vs.econgrow = Q111,
    # --------------------------- RELATIONSHIP BETWEEN GOVERNMENT AND CITIZENS
    cheating.taxes = Q180,
    gov.video.surveillance = Q196,
    gov.email.monitoring = Q197,
    gov.collecting.info = Q198,
    # --------------------------- ETHICAL VALUES ---------------------------
    terrorism = Q192,
    death.penalty = Q195,
    suicide = Q187,
    beating.wife = Q189,
    beating.children = Q190,
    # --------------------------- SOCIAL VIEWS ---------------------------
    homosexuality = Q182,
    prostitution = Q183,
    abortion = Q184,
    divorce = Q185,
    casual.sex = Q193,
    sex.before.marriage = Q186,
    # --------------------------- CAREER VALUES ---------------------------
    importance.leisure.time = Q3,
    importance.work = Q5,
    # --------------------------- IMMIGRATION ---------------------------
    imm.threatens.jobs.when.scarce = Q34,
    imm.fills.useful.jobs = Q122,
    imm.strengthens.cultural.div = Q123,
    imm.increases.crime.rate = Q124,
    imm.gives.political.asylum = Q125,
    imm.increases.terrorism.risk = Q126,
    imm.helps.poor = Q127,
    imm.increases.unemployment = Q128,
    imm.brings.social.conflict = Q129,
    imm.policy.preference = Q130
    )
```

---

## Exploratory Data Analysis

**Skimming the data**:

```{r}
myskim <- skim(wvs_subset)
myskim
```

The subset of the dataset contains `r nrow(wvs_subset)` rows and `r ncol(wvs_subset)` columns. As many of the variables are ordinal (Likert scales), the descriptive statistics generated by the skim give an accurate picture of their distribution and the characteristics. However, certain variables such as `Ethnic.Group`, `Country.and.year`, `Longitude`, and `Latitude` contain nominal, numeric data, and thus the descriptive statistics for these variables can be ignored as it is meaningless.

Most variables are numeric; only `Country` is classified as a character vector. `Longitude` and `Latitude` are missing the most data with a completion rate of `r myskim$complete_rate[2]`, but overall the mean completion rate is `r mean(myskim$complete_rate)`. This overall high completion rate signifies that missing data is not too much of an issue for the particular variables we are considering. However, it is necessary to assess whether missing data should be removed, retained, or imputed.

The `aggr` function from the VIM package can help in visualizing combinations of missing values, which will be useful for understanding what variables are missing the most data and what combinations of variables are missing the most data. Keeping in mind our subset contains `r nrow(wvs_subset)` rows, the number on the bottom-right of each plot below shows the number of "complete cases," where no data is missing.

```{r message=FALSE, warning=FALSE}
# ordinal, numeric data only
wvs_subset_ordinal <- 
  wvs_subset %>% 
  select(-Longitude,
         -Latitude,
         -Country, 
         -Ethnic.Group, 
         -Country.and.year,
         -Religion)

wvs_subset_demographic <- wvs_subset_ordinal[1:10]
wvs_subset_political <- wvs_subset_ordinal[11:16]
wvs_subset_ethical <- wvs_subset_ordinal[21:25]
wvs_subset_social <- wvs_subset_ordinal[26:31]
wvs_subset_immigration <- wvs_subset_ordinal[34:40]

par(oma=c(0,0,2,0))
plot(aggr(wvs_subset_demographic, plot = FALSE), 
     numbers = TRUE, 
     cex.axis = .5, 
     prop = FALSE,
     only.miss = TRUE)
title("Missingness in Demographic Data", outer = TRUE)

par(oma=c(0,0,2,0))
plot(aggr(wvs_subset_political, plot = FALSE), 
     numbers = TRUE, 
     cex.axis = .4, 
     prop = FALSE,
     only.miss = TRUE)
title("Missingness in Political Participation Data", outer = TRUE)

par(oma=c(0,0,2,0))
plot(aggr(wvs_subset_ethical, plot = FALSE), 
     numbers = TRUE, 
     cex.axis = .6, 
     prop = FALSE,
     only.miss = TRUE)
title("Missingness in Ethical Values Data", outer = TRUE)

par(oma=c(0,0,2,0))
plot(aggr(wvs_subset_social, plot = FALSE), 
     numbers = TRUE, 
     cex.axis = .6, 
     prop = FALSE,
     only.miss = TRUE)
title("Missingness in Social Views Data", outer = TRUE)

par(oma=c(0,0,2,0))
plot(aggr(wvs_subset_immigration, plot = FALSE), 
     numbers = TRUE, 
     cex.axis = .3, 
     prop = FALSE,
     only.miss = TRUE)
title("Missingness in Immigration Data", outer = TRUE)
```

Although the overall completion rate is relatively high (`r mean(myskim$complete_rate)`), the VIM plots reveal that missing data is an ubiquitous issue. Furthermore, our predictive modeling efforts will require complete datasets with no missing values. 

Imputing the data is not a good option. The WVS dataset contains qualitative survey data, and imputing it would be both unethical and misrepresentative. Each record contains the beliefs and experiences of a particular individual. While the mean value of beliefs on homosexuality in the dataset is `r mean(wvs_subset$homosexuality, na.rm = TRUE)` out of 10 (1 = never justifiable, 10 = always justifiable), if we were to assign this mean value to all those who chose not to answer this question, we would inaccurately be contributing to the regression toward the mean.

Retaining the data would be ideal, but our predictive modeling techniques such as decision trees require "complete cases," where all records must be present. For this reason, we are opting to remove records with missing values. The impact of our decision can be seen below: `wvs_subset_ordinal` had `r nrow(wvs_subset_ordinal)` rows before cleaning, and ...

```{r}
wvs_subset_ordinal <- na.omit(wvs_subset_ordinal)
```

... after cleaning `wvs_subset_ordinal` has `r nrow(wvs_subset_ordinal)` rows.

Although this reduces the size of the dataset by a considerable portion, it is the best option based on our inability to impute nor keep the data.

**Correlations in the data**:

```{r}
# further subsetting to topics of highest interest
wvs_subset_correlations <- 
  wvs_subset_ordinal %>% 
  select(
    # --------------------------- DEMOGRAPHICS
    Settlement.type,
    Town.size,
    Age,
    Income.Group,
    Immigrant,
    Marital.Status,
    Education,
    Number.Children,
    Happiness,
    Health,
    # ------------------- ETHICAL VALUES / SOCIAL VIEWS -------------------
    terrorism,
    death.penalty,
    suicide,
    beating.wife,
    beating.children,
    homosexuality,
    prostitution,
    abortion,
    divorce,
    casual.sex,
    sex.before.marriage,
    # --------------------------- IMMIGRATION ---------------------------
    imm.threatens.jobs.when.scarce,
    imm.fills.useful.jobs,
    imm.strengthens.cultural.div,
    imm.increases.crime.rate,
    imm.gives.political.asylum,
    imm.increases.terrorism.risk,
    imm.helps.poor,
    imm.increases.unemployment,
    imm.brings.social.conflict,
    imm.policy.preference,
  )

# assessing correlations between ordinal, numeric data
# only including our areas of focus: demographics, religion, ethics, immigration
# only including data where pairwise observations are complete (i.e., not including missing data)
cor <- cor(wvs_subset_correlations, use = "pairwise.complete.obs")
corrplot(cor, 
         method = "circle", 
         insig = "blank", 
         diag = FALSE,
         tl.cex = 0.5) %>% 
    corrRect(name = c('Settlement.type', 
                      'terrorism', 
                      'imm.threatens.jobs.when.scarce',
                      'imm.policy.preference'))
```

The visualization of a correlation matrix above indicates the extent to which each variable, compared to one another, correlates or relates. Darker blue colors represent stronger positive correlations, and darker orange colors represent stronger negative correlations. For visualization purposes, we have bucketed the variables into four groups, but it should be noted that by doing so we risk imposing structure where it does not, or should not, exist: 

- The first bucket contains variables relating to demographics
- The second bucket relates to ethical values / social views
- The third bucket relates to immigration / beliefs about immigrant populations

While correlation analysis is typically reserved for numeric variables, we felt it would be a useful exploratory analysis to get an approximate feel for the types of relationships that may inform our EDA and later, our classification decision tree algorithms. Our options for correlation with numeric data were limited. Our numeric values are limited to demographic information rather than values, as values are measured on a Likert scale. Furthermore, there is a growing body of research in support of using ordinal variables as if they were continuous or factor variables for correlation or other analysis (such as factor analysis) is acceptable (Robitzsch, 2020). Robitzsch, a German researcher, asserts that there is no "correct" modeling strategy, and that ordinal variables may be just as well suited for non-ordinal models as they are for ordinal ones. Treating ordinal variables as continuous, and using Pearson correlations to show relationships between them, may result in less bias than treating the ordinal variables as ordinal. 

Although this visualization of a correlation matrix reveals many interesting insights into how our variables correlate, the following are just a couple of key insights relating to the focus of our project — religion, immigration, and ethics:

- Some correlations occur between variables in the first bucket. For example, `Settlement.type` and `Town size` are strongly correlated; rural areas have few residents, urban areas have many. Similarly, `Marital Status` correlates with `Age` and `Number.Children`; `Income.Group` correlates with `Education`; `Happiness` correlates with `Health`. These are all to be expected.
- Many correlations occur between variables in the second bucket. Likely, this is an indication of liberal versus conservative thought. For example, those who hold liberal/conservative views on `homosexuality` are also likely to hold respectively liberal/conservative views on `abortion`.
- Some correlations occur between variables in the third bucket.For example, those who feel immigration does (or does not) increase crime rates also feel immigration does (or does not) increase social conflict.

Our correlation matrix has confirmed that we humans are not blind in our values and beliefs. We seek order and consistency. If we feel one way about one thing, we will likely feel similarly about another, related thing. That said, we are complex, and thus the correlation matrix shows a degree of inconsistency, too. 

Having understood what correlates with what, the next step in our exploratory data analysis will be to explore where clusters form. Using geospatial visualization techniques, we will assess how views on religion, immigration, ethics are spread across the globe.

**Visualizing the data**:

```{r}
# Creates colored points (gradient) based on particular categorical numeric values in dataframe.
# For display in a map (see examples below)
map_color_range <- function(tibl, col, 
                            min_color = "green", max_color = "red",
                            shape = 1, size = 0.4) {
  min_value <- min(col, na.rm = TRUE)
  max_value <- max(col, na.rm = TRUE)
  # vector of color gradients with a length of `max_value`
  colors <- colorRampPalette(c(min_color, max_color))(max_value)
  # executes points() for each value between min and max
  for (i in min_value:max_value) {
    tibl_subset_col <- 
      tibl %>% 
      filter(col == i) 
    points(tibl_subset_col$Longitude, 
           tibl_subset_col$Latitude, 
           col = alpha(colors[i], 0.1), # colors from color ramp above w/ transparency
           pch = shape, 
           cex = size)
  }
}

map("world")
title(main = "Homosexuality is justifiable (green) vs. never justifiable (red)")
map_color_range(wvs_subset, wvs_subset$homosexuality, 
                min_color = "red", max_color = "green")
```

Geospatial data can reveal many interesting trends. For example, we can visualize beliefs on `homosexuality`, where greener colors represent those who believe homosexuality is "always justifiable" and redder colors represent those who believe homosexuality is "never justifiable." Eastern Europe and parts of Africa appear most opposed to homosexuality, whereas North America appear most tolerant.

```{r}
map("world")
title(main = "Immigration does (green) or does not (red) increase cultural diversity")
map_color_range(wvs_subset, wvs_subset$imm.strengthens.cultural.div, 
                min_color = "red", max_color = "green")

map("world")
title(main = "Immigration does (red) or does not (green) increase risk of terrorism")
map_color_range(wvs_subset, wvs_subset$imm.increases.terrorism.risk, 
                min_color = "green", max_color = "red")
```

The two visualizations above, when compared with one another, reveal an interesting dichotomy in beliefs on immigration. On the one hand, people across the globe generally say immigration strengthens cultural diversity. On the other hand, people across the globe also generally say immigration increases the risk of terrorism.

So the question arises, do people generally support or not support immigration? What this visualization tells us is that humans have conflicting beliefs. It may just be the case that we humans can hold conflicting beliefs. In the field of psychology, this is referred to as "cognitive dissonance." As a classic example, meat-eaters may believe that harming animals is a bad thing, and yet the act of eating meat contributes to the suffering of animals. Our team feels that beliefs on immigration may be similar. Humans across the globe generally like the idea of cultural diversity, and yet when it comes to living alongside people of different cultures, they may exhibit fearful or cautious behaviors that suggest otherwise.

The human experience is a many-faceted. Now that we have explored what correlates with what, and where clusters form, our next step will be to apply decision trees as a predictive modeling technique to understand more precisely what variables are most indicative of certain values.

---

## Predictive Modeling Techniques and Interpretations

In each of our decision trees, 80% of the data will randomly be placed into a training set and 20% into a test set. The training sets will be used to create the decision trees, and the test sets will be used to test the resulting model's accuracy. The `rpart` function will build a model separating the outcome variable into distinct groups, and the model will determine which variables are deemed most important for classifying the outcome variable. Node 1 will always serve as the best classifier for splitting the data into two groups, and the following nodes serve as the next best classifiers for splitting the data into groups. This process repeats recursively until no further splitting is possible or necessary at which point a decision has been reached.

The major question to ask of a decision tree is whether the model is too complex, too specific, or too sensitive. In other words, did the recursive process of splitting the data into groups and subgroups continue on for too long. When should the splitting stop? For each of our business questions, we will apply the decision tree to our test data and then to our train data to assess and explain the model's accuracy.

##### Business Question #1: Classifying Immigrants versus Nonimmigrants

In this first decision tree, we seek to understand what variables are most indicative of whether or not someone is an immigrant. A human posed with this same question might list a number of factors such as education, income, or number of children, but these would be biased. A decision tree, on the other hand, will assess all relationships between all variables in the dataset to produce a simple rule-based model.

```{r}
immigrant_tree <- 
  wvs_subset %>% 
  select(-Longitude, -Latitude, -Country.and.year, -Ethnic.Group, -Country)

# partition prep
set.seed(888)
# remove NAs
immigrant_tree <- na.omit(immigrant_tree)

# partition
immigrant_tree_partition <- createDataPartition(y = immigrant_tree$Immigrant, p = 0.8, list = FALSE)
immigrant_tree_train <- na.omit(immigrant_tree[immigrant_tree_partition, ])
immigrant_tree_test <- immigrant_tree[-immigrant_tree_partition, ]

# building the classification tree with rpart
immigrant_tree_model <- rpart(formula = Immigrant ~ .,
                              data = immigrant_tree_train,
                              method = "class")

# Visualize the decision tree with rpart.plot
fancyRpartPlot(immigrant_tree_model, 
               main = "Decision Tree #1: Nonimmigrant or Immigrant\n-- a question of political participation, immigration beliefs, and age --",
               caption = "Nonimmigrant = 1, Immigrant = 2",
               type = 5)
```

The decision tree boiled all variables down to just several splits. The path to being classified as an "Immigrant" (2) can be read as follows:

- Do you rarely vote locally, or are you not allowed to vote locally? If so, you may be an immigrant.
- Do you disagree strongly that immigration threatens jobs? If so, you may be an immigrant.
- Are you 24 years or older? If so, you may be an immigrant.
- Do you rarely vote nationally, or are you not allowed to vote nationally? If so, you are an immigrant.
- If the answer is no to any of the above questions, you are not an immigrant.

What is interesting about these variables is that the first and last relate to political participation, the second relates to a belief about immigration, and the third relates to a demographic. The model has incorporated a convincing spread of variables to come up with its rule-based output. We can next test for the accuracy of the model using the test set.

```{r}
#Testing the model
pred_immigrant_tree_model <- predict(object = immigrant_tree_model, immigrant_tree_test[-5], type = "class")
immigrant_tree_model_matrix <- table(immigrant_tree_test$Immigrant, pred_immigrant_tree_model)
confusionMatrix(immigrant_tree_model_matrix)
```

This model is highly accurate (94.77%)! This shows that the model does a good job on unseen data. 

##### Business Question #2: Predicting whether an individual views teaching tolerance as important based on their beliefs regarding immigration 

In this second decision tree, we seek to understand attitudes towards immigration in relation to a key value: tolerance. By setting up a decision tree with two questions on immigration that are in some ways in conflict with one another, we sought to answer the question, "how do those who believe tolerance is an important teaching for children feel about immigrants in their immediate lives, and more broadly, as a threat to security?" The two variables employed to answer this question are to the effect of, "does immigration increase the risk of terrorism?" and "would you like to have immigrants as neighbors?"

The following decision tree predicts the importance of teaching children tolerance based on two viewpoints on immigration: friendliness towards immigrants as neighbors and whether immigration increases the risk of terrorism. A decision tree was chosen as the mode of analysis because it provides a thorough answer to the question of whether those who emphasize tolerance to their children display tolerance towards immigrants in their home country.

```{r}
# sub-setting data
tree_subset <- 
  wvs %>% 
  select(Immigrants.As.Neighbors = Q21, # 1 = mentioned they wouldn't like this,                                             2 = didn't mention
        Tolerance.Important.For.Children = Q12, # 1 = mentioned to be important,                                                     2 = didn't mention
        Immigration.Increases.Terrorism.Risk = Q126) # 2 = agree, 1 = unsure, 0 = disagree
# skim
tree_skim <- skim(tree_subset)
# prop tables
prop.table(table(tree_subset$Immigrants.As.Neighbors)) # 1 = mentioned they wouldn't like this,                                                  2 = didn't mention
prop.table(table(tree_subset$Tolerance.Important.For.Children)) # 1 = mentioned to be important,                                                      2 = didn't mention
prop.table(table(tree_subset$Immigration.Increases.Terrorism.Risk)) # 2 = agree, 1 = unsure,                                                                     0 = disagree
# partition prep
set.seed(1000)
# remove NAs
tree_subset <- na.omit(tree_subset)
# partition
data_partition<-createDataPartition(y=tree_subset$Tolerance.Important.For.Children, p=0.8, list=FALSE)
train <- na.omit(tree_subset[data_partition,])
test <- tree_subset[-data_partition,]
# checking partition and NA omits
nrow(tree_subset)
nrow(train)
nrow(test)
head(train)
# tree plotting
library(rpart)
tolerance_tree <- rpart(formula = Tolerance.Important.For.Children ~., 
                 data = train,
                 method = 'class',
                #minsplit=2,
                cp=-1,
                 #minbucket = 2,
                #parms = list(split = 'gini')
                )

# tree visualization
library(rattle)
library(rpart.plot)
fancyRpartPlot(tolerance_tree, main = "Decision Tree #2: Importance of Teaching Tolerance,\nBased on Beliefs about Immigration", caption = "Tolerance is Important (1) versus Nonimportant (2)", type = 5)
# Immigrants as neighbors: 1 = wouldn't like  2 = didn't mention
                           #0.2199076           0.7800924
# Tolerance important for children: 1 = yes   2 = didn't mention 
                                  # 0.6282186    0.3717814
# Immigration increases terrorism risk: # 2 = agree, 1 = unsure, 0 = disagree
                                        # 0.4558246,  0.2214990,  0.3226764

# Variable of importance
tolerance_tree$variable.importance # immigrants as neighbors

# Test variable prediction
tree_pred<-predict(tolerance_tree,newdata=test,
                   type = 'class',
                   prob = TRUE)

# Measure model performance
library(caret)
table(tree_pred, test$Tolerance.Important.For.Children)
confusionMatrix(as.factor(tree_pred), as.factor(test$Tolerance.Important.For.Children))
```

**Findings**:

It may be assumed that immigration viewpoints would be a strong predictor of importance of teaching children tolerance. If people view immigrants as an important aspect of their home home country culture and communities, we would likely describe them as tolerant, among other things. Therefore, analyzing the results of the decision tree with one positive question about immigration and one negative, it's interesting to see the results on the leaves, which all land close the 50/50 line, with tolerance being important teaching for children holding the majority in each leaf. The decision tree is 63% accurate. Decision trees tend to over fit accuracy without extensive preparation to the training and test partitions.

**Tolerance Importance: Leaf Interpretation (L-R)**:

- The vast majority of respondents - *78%* - are neutral of positive about the prospect of having neighbors that have immigrated from other countries.
- Of these 78%, 34% of respondents are neutral or positive about immigrants as neighbors, and responded that immigration increases the risk of terrorism. I'll classify this group as somewhat supportive of immigration (open to immigration in their immediate community and wary of immigration in their country). Within this leaf, 65% believe in the importance of teaching children tolerance (the highest among the leaves).
- 27% are neutral or positive about immigrants as neighbors, and disagree that immigration increases the risk of terrorism. I'll classify this group as most open to immigration. Within this leaf, 64% believe in the importance of teaching tolerance to children.
- 18% are neutral or positive about immigrants as neighbors, and unsure about the risk they pose to terrorism. This group is neutral. 
- Within the much smaller cohort that responded that they wouldn't like having immigrants as neighbors - *22%* - 10% are neutral or disagree with the question of whether immigiration increases the risk of terrorism. I'll include this group in the somewhat supportive of immigration category. This group also believes in teaching tolerance to children, though by a slim margin of about 6%. 
- In contrast, 12% of respondents agree that immgiration increases the risk of terrorism, and responded negatively to the idea of being neighbors with immigrants. This group is classified as the most unfavorable to immigration. Still, 59% of this group believes in teaching tolerance to children.

Based on this classification system, respondents fall into the following categories regarding immigration viewpoints: 27% are very supportive, 44% are somewhat supportive, 18% are neutral or unsure, and 12% are not supportive. The question of whether those who emphasize tolerance to their children display tolerance towards immigrants in their home country is answered; every category classified by immigration viewpoints, including those with negative attitudes, responded that they believe in teaching tolerance to children.

##### Business Question #3: Classifying Religious Denominations

In this third decision tree, we seek to understand what variables are most indicative of religious denomination. A human may observe the foods someone eats or the rituals someone practices to determine this, but it will be interested to feed variables in our dataset to a decision tree to understand which factors are found to be most important in this classification task.

```{r}
religion_tree <-
  wvs_subset %>% 
  select(-Longitude, -Latitude, -Country.and.year, -Ethnic.Group, -Country)

# partition prep
set.seed(888)
# remove NAs
religion_tree <- na.omit(religion_tree)

# partition
religion_tree_partition <- createDataPartition(y = religion_tree$Religion, p = 0.8, list = FALSE)
religion_tree_train <- na.omit(religion_tree[religion_tree_partition, ])
religion_tree_test <- religion_tree[-religion_tree_partition, ]

# building the classification tree with rpart
religion_tree_model <- rpart(formula = Religion ~ .,
                             data = religion_tree_train,
                             method = "class")

# Visualize the decision tree with rpart.plot
fancyRpartPlot(religion_tree_model, 
               main = "Decision Tree #3: Religious Denomination\n-- a question social views, ethical values, and town size --",
               caption = "0 = No religion, 1 = Roman Catholic, 5 = Muslim",
               type = 5)
```

The decision tree boiled all variables down to just several splits. It even chose to take many of the possible religions out of the question, likely because there were too few of each to include. The path to being classified as "Roman Catholic" (1) can be read as follows:

- Do you believe sex before marriage is justifiable to some extent? If so, you may be Roman Catholic.
- Do you live in a town/city with less than 20,000 residents? If so, you may be Roman Catholic.
- Do you believe homosexuality is not justifiable to some extent? If so, you may be Roman Catholic.

Here, the first and last variable considered by the model relates to social views / ethical values, while the second variable relates to a demographic. Again, the model has incorporated just what it thinks to be of utmost importance to come up with its rule-based output. Other interpretations we can make based on this model are:

- 63% of those in the training set who think sex before marriage is justifiable to some extent either do not have a religion or are Roman Catholic, the rest are Muslim.
- The classification of Muslims does not include their town size or views on homosexuality. The model would say that beliefs on sex before marriage, alone, is enough to classify whether or not someone is Muslim.

Already, this model seems slightly problematic. There may be people who are not Muslim who similarly believe sex before marriage is never justifiable, but the model at present would classify all people who share this belief as Muslim. As a next step, we can test the accuracy of the model by feeding it the test set.

```{r}
#Testing the model
pred_religion_tree_model <- predict(object = religion_tree_model, religion_tree_test[-6], type = "class")
religion_tree_model_matrix <- table(religion_tree_test$Religion, pred_religion_tree_model)
confusionMatrix(religion_tree_model_matrix)
```

This model is highly inaccurate (38.45%)! This shows that the model does a poor job on unseen data. A possible reason for this is the skewness in the distributions of data on different religious denominations. A recommended next step would be to employ random sampling techniques on the dataset to normalize the distributions of each religious denomination prior to creating training and test splits.

##### Business Question #4: Predicting Confidence in Elections

In this fourth and final decision tree, we seek to answer the question of whether those who have faith in elections also have faith in the courts and the United Nations (UN). Elections operate as the bedrock of democracy, but in countries without democracy, elections are often futile systems (or at least perceived to be). 

Taking a quick step back to exploratory visualizations, it would be interesting to see, globally, how confident people are about their countrys' elections.

```{r}
map("world")
title(main = "Confident (green) vs. not confident (red) in election process")
map_color_range(wvs_subset, wvs_subset$confidence.elections, 
                min_color = "green", max_color = "red")
```

It appears overall we humans do not trust elections! Therefore, it's worthwhile and interesting to assess the relationship, if any, between confidence in elections, courts, and the UN. A decision tree was chosen for this reason, and because these questions were answered on the same Likert scale, making for a clearer classification from the algorithm.

```{r}
tree_four <- 
  wvs_subset %>% 
  select(-Longitude, -Latitude, -Country.and.year, -Ethnic.Group, -Country)

# partition prep
set.seed(1200)

# remove NAs
tree_four <- na.omit(tree_four)

# partition
four_partition <- createDataPartition(y = tree_four$confidence.elections, p = 0.8, list = FALSE)
four_train <- na.omit(tree_four[four_partition, ])
four_test <- tree_four[-four_partition, ]

# building the classification tree with rpart
four_model <- rpart(formula = confidence.elections ~ .,
                             data = four_train,
                             method = "class")

# Visualize the decision tree with rpart.plot
fancyRpartPlot(four_model, 
               main = "Decision Tree #4: Predicting Confidence in Elections",
               caption = "Confidence Level:\n1 = A great deal, 2 = Quite a lot, 3 = Not very much, 4 = Not a lot",
               type = 5)
```

- The response scale is the same among each of the variables included in this decision tree. Confidence in courts and in the United Nations (UN) are used to predict confidence in elections. 

**Confidence in Elections: Leaf Interpretation (L-R)**:

- High and low confidence in the courts is split almost 50/50, with 51% of respondents reporting at least quite a bit of confidence in the court, and 49% of respondents reporting less than or equal to not very much confidence in the court system. 
- 5% of respondents have a great deal of confidence in the UN, courts, and elections. This means that only 1 in 20 people have a lot of confidence in all 3 bodies in their country.
- 9% of respondents have a great deal of confidence in courts, and quite a lot of confidence in the UN and elections.
- 37% of respondents have quite a lot of confidence in courts and elections.
- 32% of respondents have not very much confidence in courts, or elections.
- 17% of respondents have not a lot of confidence in courts or elections.

In summary, results indicate that half of respondents have little trust in public systems like the courts, elections, and the UN, while the other half have moderate-to-high levels of trust in each. A mere 5% of respondents rated all 3 institutions with the highest level of trust, validating the popular culture idea that there is widespread public skepticism concerning political systems. 

Next, we will assess the model's accuracy by feeding it the test set.

```{r}
# Variable of importance
four_model$variable.importance # confidence in courts

# Test variable prediction
four_pred<-predict(four_model,newdata=four_test,
                   type = 'class',
                   prob = TRUE)

# Measure model performance
library(caret)
table(four_pred, four_test$confidence.elections)
confusionMatrix(as.factor(four_pred), as.factor(four_test$confidence.elections))
```

Model accuracy is 54%. Decision trees do not consider the interdependence of variables when calculating a split, so using a random forest instead of one decision tree results in a more accurate model. Another way to improve accuracy is k-fold cross validation. For the purposes of this business question, the model and it's accuracy suffice.

---

## Interpretations

**Summary**:

Our initial exploratory data analysis and subsequent findings from decision trees have begun to supply answers on each of our decision trees, which we will summarize below:

- Immigration-related
-- #1: Predicting whether or not an individual is an immigrant
-- #2: Predicting whether an individual views teaching tolerance as important based on their beliefs regarding immigration 

Our geospatial visualizations revealed that views on immigration are contradictory. People across the globe value cultural diversity and yet are simultaneously fearful of immigrants. In terms of what it means to be an immigrant, our first decision tree found that it is a combination of demographic factors, paired with political participation, paired with ethical values. Our second decision tree found that people generally value teaching tolerance in their households irrespective of their beliefs regarding immigration, which is somewhat counter-intuitive as it would be natural to think that views on tolerance and views on immigration correlate.

- #3: Predicting the religious denomination of an individual

Our third decision tree brought together an interesting pairing of ethical values and demographic factors as determinants of religious denomination. Although the model accuracy was low, there is likely truth in the statement that religious denomination goes hand-in-hand with ethical values. Demographic factors such as town size, too, must have an affect on religious denomination, as smaller towns may be more communal and/or their inhabitants more reliant upon one another and thus more homogeneous in terms of religious denomination, whereas larger, urban areas might be more agnostic due to their diversity.

- #4: Predicting an individual's confidence in elections

Our final decision tree and accompanying geospatial visualization revealed the world, overall, is not very confident in governmental organizations and structures. However, those who believe in courts are more likely to also believe in elections, and those who believe in both are more likely to believe in the UN, too. This is another example of humans exhibiting consistent identity traits, where those who identify as supporters of governmental structures or organizations are likely to consistently identify as supporters of all-things government. On the other hand, those lacking trust in governmental structures or organizations will generally lack trust across the board when it comes to such topics.

**Final Thoughts**:

Humans are complex. We are moderately to extremely consistent on some things (e.g., beliefs on governmental organizations, religion, ethics), but then we are moderately to extremely inconsistent on other things (e.g., immigration). We set out to explore the various facets of the human experience, and we now culminate with a deeper understanding of what makes a human a human.

## Recommendations & Conclusions

- Next, we plan to integrate feedback from Professor Hodeghatta to our final report and presentation.
- We will be selecting aspects of our analysis to include in a PowerPoint presentation for classmates next Wednesday, which will require zeroing in on a select few analyses that are visually-focused, namely graphs and decision tress.
- Future analysis would touch on those issues that we were not able to cover. These included the relationship between education completed (Q275) and the importance of leisure time (Q3); the relationship between income (Q288), happiness (Q46), and health (Q47); and the relationship between marital status (Q273) and political engagement (Q209 - Q222).
- Future analysis would include more advanced predictive modeling techniques, starting with random forests. We cannot know what we have not yet studied, but we would hope to find other useful modeling techniques beyond decision trees that pair well with ordinal (Likert) data!

## Works Cited
Robitzsch, A. (2020). *Why Ordinal Variables Can (Almost) Always Be Treated as Continuous Variables: Clarifying Assumptions of Robust Continuous and Ordinal Factor Analysis Estimation Methods.* Frontiers. https://www.frontiersin.org/articles/10.3389/feduc.2020.589965/full